---
title: "Morphological Data Analysis"
output: html_notebook
---
Author: Indiana Riley

#***Data Standardization***
#1. Setup


```{r}
#set working directory
setwd("~/Data Analysis")
#import raw data
raw_dat <- read.csv("data/Whole_Fish_Raw_Measurements.csv", header = TRUE)
#import location numbers
loc <- read.csv("data/Whole_Fish_Locations.csv", header = TRUE)
#load required packages
library(tidyverse)
library(dplyr)
library(ggplot2)
```

#2. Cleaning dataset


```{r}
raw_dat <- raw_dat%>%
  filter(Species != "?") %>% # removes all species with unknown identification
  filter(Sample_ID != "E30") %>% # removing damaged specimen
  mutate(LogSL = log(SL)) # adding log-transformed standard length
#adding site numbers to the main dataset
raw_dat <- left_join(raw_dat, loc, by = "Location")

#making character strings of values for use in for loops
all_headers <- colnames(raw_dat)
variables <- all_headers[-c(1:2, 5:13, 15, 29:30)]
meristics <- all_headers[-c(1:2, 5, 7, 9, 15:30)]
stand_vars <- variables[-c(1:2, 17:18)]
loop_vars <- all_headers[c(16:28)]

#log transforming all variables
stand_dat <- raw_dat %>%
  mutate_at(stand_vars, funs(log(.)))
write.csv(stand_dat, file = "~/Data Analysis/results/standardizing data/log_standardized_data.csv", row.names = F)

```

#3. Extracting residual values

This for loop goes through every log-standardized morphological measurement and:
a) fits a linear model between the measurement and standard length 
b)prints graphs to assess residuals, normality and homogeneity of variance
c) extracts the linear model residuals, converts them to a data frame, and saves them as a csv

You can then combine all these separate files in excel and import them back into R ( "master_sheet_residuals.csv")

```{r}

for (i in loop_vars){
  print(i)
  
  temp_model <- lm(paste('stand_dat$', i, '~ stand_dat$SL', sep = ""), na.action = na.exclude)
  
  jpeg(paste('~/Data Analysis/results/standardizing data/model_', i, '.jpeg'), width = 1000, height = 500)
  par(mfrow = c(1,3))
  plot(temp_model, which = 1)
  hist(temp_model$residuals) 
  plot(temp_model, which = 2) 
  dev.off()
  
  model_residuals <-resid(temp_model)
  
  d <- as.data.frame(model_residuals)
  
  colnames(d) <- as.character(i)
  
  write.csv(d, file = paste("~/Data Analysis/results/standardizing data/", i,"_residuals.csv", sep = ""), row.names = FALSE)
  
}

all_reg <- read.csv("~/Data Analysis/results/standardizing data/master_sheet_residuals.csv", header = T)
```

#***Principal Component Analysis***

#Documentation for creating pca dataset (do not rerun)


```{r}
# Reading in three files - contain only values for definite species
#   fish_dat contains the transformed values for the morphometrics 
#     (residuals of log-measurements on log SL)
#   mer_dat contains the meristics for the same set of samples 
#     (and log standardized data for morphometrics that will be cut off)
#   loc contains numbers for each location (for use in manipulating graph colours)
#fish_dat <- read.csv("results/residuals/master_sheet_residuals.csv", header = TRUE)
#mer_dat <- read.csv("results/log_standardized_data.csv", header = TRUE)
#loc <- read.csv("data/Whole_Fish_Locations.csv", header = TRUE)

# converting NA values for morphometrics to 0
#fish_dat$LJL[is.na(fish_dat$LJL)] = 0

#converting NA values for meristics to pooled means and cutting off unnecessary data
#mer_dat2 <- mer_dat %>%
  #select(2:6, 8, 10:13) %>%
  #group_by(Species, Location) %>%
  #mutate_at(9:10, funs(ifelse(is.na(.), mean(., na.rm = TRUE),.)))

# cutting off more unnecessary data
#mer_dat3 <- mer_dat2 %>%
  #ungroup() %>%
  #select(1, 5:10)

# joining the three datasets and saving as a csv to relaod next time
#all_pca_data <- left_join(fish_dat, mer_dat3, by = "Sample_ID")
#all_pca_data <- left_join(all_pca_data, loc, by = "Location")
#write.csv(all_pca_data, file = "results/PCA/pca_data.csv", row.names = FALSE)

```

#1. Setup


```{r}
# this dataset contains all the residual-standardized morphometrics and unstandardized meristics for each sample (See above)
dat_pca <- read.csv("~/2021-Docs/2021 Uni/Publishing/All R Work/results/PCA/pca_data.csv", header = TRUE)

#load all packages required
library(tidyverse)
library(dplyr)
library(ggplot2)
library(factoextra)
library(ggbiplot)

```

#2. PCA

```{r}

#creating individual dataframes with only morphometric and meristic variables
morph.df <- dat_pca[c(5:17)]
mer.df <- dat_pca[c(18:23)]
mer.df <- mer.df %>% 
  dplyr::rename(
    DOR = DOR_XRAY,
    ANA = ANA_XRAY
  )
#PCA - morphometric data
morph.pca <- prcomp(morph.df, center = TRUE, scale = TRUE)
summary(morph.pr)
res.var <- get_pca_var(morph.pr)
res.var$coord
res.var$contrib
res.var$cos2
eig.val <- get_eigenvalue(morph.pr)
#PCA - meristic data
mer.pca <- prcomp(mer.df, center = TRUE, scale = TRUE)
summary(mer.pr)
res.var <- get_pca_var(mer.pr)
res.var$coord
res.var$contrib
res.var$cos2

```

#3. Plotting PCA

```{r}
#loading required package for ggplot to interpret PCA objects
library(ggfortify)

#plotting morphometric PCA
morph_pca_graph <- autoplot(morph.pca, loadings = TRUE, loadings.label = TRUE, data = dat_pca, 
         colour = 'Species', shape = 'Species', fill = 'Species', size = 1.75,
         loadings.label.colour = "gray20", loadings.colour = "gray20",
         loadings.label.size = 3.5,loadings.label.fontface = "bold",
         loadings.label.repel = TRUE,
         loadings.label.hjust = 1) +
  scale_shape_manual(name = "Species", values=c(21, 24)) +
  scale_color_manual(name = "Species", values = c("darkorange1", "deepskyblue1")) +
  theme_classic() +
  theme(legend.position = "bottom",
        axis.title = element_text(size = 12),
        axis.text = element_text(colour = "black", size = 10),
        legend.title = element_text(size = 13),
        legend.text = element_text(face = "italic", size = 12))

#plotting meristic PCA
mer_pca_graph <- autoplot(mer.pca, loadings = TRUE, loadings.label = TRUE, data = dat_pca, 
         colour = 'Species', shape = 'Species', fill = 'Species', size = 1.75,
         loadings.label.colour = "gray20", loadings.colour = "gray20",
         loadings.label.size = 3.5,loadings.label.fontface = "bold",
         loadings.label.repel = TRUE,
         loadings.label.hjust = 1) +
  scale_shape_manual(name = "Species", values=c(21, 24)) +
  scale_color_manual(name = "Species", values = c("darkorange1", "deepskyblue1")) +
  theme_classic() +
  theme(legend.position = "bottom",
        axis.title = element_text(size = 12),
        axis.text = element_text(colour = "black", size = 10),
        legend.title = element_text(size = 13),
        legend.text = element_text(face = "italic", size = 12))

#saving plots
ggsave("~/Data Analysis/results/PCA/morph_pca_graph.jpg", morph_pca_graph)
ggsave("~/Data Analysis/results/PCA/mer_pca_graph.jpg", mer_pca_graph)

```

#***Meristics***

#Documentation for creating graph dataset (do not rerun)

```{r}
#raw_data <- read.csv("data/Whole_Fish_ID_for_Museum.csv", header = T)
#locations <- read.csv("data/Whole_Fish_Locations.csv", header = T)
#raw_data <- raw_data %>% 
#  filter(!is.na(Species),
#         Sample_ID != "E28",
#         Sample_ID != "E30")
#raw_data <- left_join(raw_data, locations, by = "Location")
#grouped_data <- raw_data %>% 
#  group_by(Location_Num, Species)
#write.csv(raw_data, file = "results/meristics/grouped_for_histograms.csv", row.names = FALSE)

```

#1. Setup

```{r}
#loading dataset
grouped_for_histograms <- read.csv(file = "~/2021-Docs/2021 Uni/Publishing/All R Work/results/meristics/grouped_for_histograms.csv", header = T)

#loading packages
library(ggplot2)
library(dplyr)
library(tidyverse)

# creating a theme for use w all histograms
#   NOTE: In future, only use strip.text.y.left if it has moved to left side, otherwise just use strip.text.y
theme_myhist <- theme(panel.grid = element_blank(),
                      panel.background = element_blank(),
                      axis.title.y = element_blank(),
                      axis.text.y = element_blank(),
                      axis.ticks.y = element_blank(),
                      legend.position = "bottom",
                      strip.placement = "inside",
                      strip.text.y.left = element_text(angle=0, hjust=1),  
                      strip.background = element_rect(colour = "white", fill = "white"))

```

#2. Cleaning dataset

```{r}
# reordering the locations according to site number for use in plotting
graph_data <- grouped_for_histograms %>% 
  mutate(LocSp = fct_reorder(LocSp, Location_Num),
         Location = fct_reorder(Location, Location_Num)) 

# removing unknown species from dataset
graph_data_2 <- graph_data %>% 
  filter(Species != "Unknown")
graph_data_2 <- graph_data_2 %>% 
  mutate(LocSp2 = fct_reorder(LocSp2, Location_Num)) 

```

#3. Creating basic histograms

```{r}
RGR1plot <- ggplot(graph_data_2, aes(x=RGR1, fill = Species)) +
  geom_histogram(bins = 12, position = "identity", alpha = 0.5, color = "black") +
  scale_fill_manual(values = c("darkorange1", "deepskyblue1", "darkorchid1"))+
  facet_grid(LocSp2~., switch = "y")+ 
  labs(x="First Arch Gill Rakers") +
  theme_myhist

RGR2plot<- ggplot(graph_data_2, aes(x=RGR2, fill = Species)) +
  geom_histogram(bins = 12, position = "identity", alpha = 0.5, color = "black") +
  scale_fill_manual(values = c("darkorange1", "deepskyblue1", "darkorchid1"))+
  facet_grid(LocSp2~., switch = "y")+ 
  labs(x="Second Arch Gill Rakers") +
  theme_myhist 

VERTplot <- ggplot(graph_data_2, aes(x=VERT, fill = Species)) +
  geom_histogram(bins = 5, position = "identity", alpha = 0.5, color = "black") +
  scale_fill_manual(values = c("darkorange1", "deepskyblue1", "darkorchid1"))+
  facet_grid(LocSp2~., switch = "y")+ 
  labs(x="Vertebrae") +
  theme_myhist

Sizeplot <- ggplot(graph_data_2, aes(x=SL, fill = Species)) +
  geom_histogram(binwidth = 5, position = "identity", alpha = 0.5, color = "black") +
  scale_fill_manual(values = c("darkorange1", "deepskyblue1", "darkorchid1"))+
  facet_grid(LocSp2~., switch = "y")+ 
  labs(x="Standard Length (mm)") +
  theme_myhist

DORplot <- ggplot(graph_data_2, aes(x=DOR_XRAY, fill = Species)) +
  geom_histogram(bins = 3, position = "identity", alpha = 0.5, color = "black") +
  scale_fill_manual(values = c("darkorange1", "deepskyblue1", "darkorchid1"))+
  facet_grid(LocSp2~., switch = "y")+ 
  labs(x="Dorsal Fin Rays") +
  theme_myhist

```

#4. Creating dataframes to plot n samples and mean onto histograms

```{r}
#Summary data for reference
summary <- graph_data %>% 
  group_by(LocSp) %>% 
  summarise(meanSL = mean(SL),
            nSL = sum(!is.na(SL)),
            meanRGR1 = mean(RGR1, na.rm = T),
            nRGR1 = sum(!is.na(RGR1)),
            meanRGR2 = mean(RGR2, na.rm = T),
            nRGR2 = sum(!is.na(RGR2)),
            meanVERT = mean(VERT),
            nVERT = sum(!is.na(VERT)),
            meanDOR = mean(DOR_XRAY),
            nDOR = sum(!is.na(DOR_XRAY)))

#Creating dataframe to use when labelling facets
prac_labels <- data.frame(LocSp2 = c("FA", "NBA", "NBM", "EA", "EM", "VM", "SAM", "PM"), 
                          label = c("NSW", "NSW", "NSW", "NSW", "NSW", "VIC", "SA", "WA"),
                          Species = c("australis", "australis", "melanochir", "australis", "melanochir", "melanochir", "melanochir", "melanochir"),
                          nSL = c("n = 10", "n = 9", "n = 1", "n = 11", "n = 11", "n = 28", "n = 17", "n = 24"),
                          nRGR1 = c("n = 10", "n = 9", "n = 1", "n = 9", "n = 7", "n = 28", "n = 17", "n = 24"),
                          nRGR2 = c("n = 10", "n = 9", "n = 1", "n = 9", "n = 8", "n = 28", "n = 17", "n = 24"),
                          nVERT = c("n = 10", "n = 9", "n = 1", "n = 11", "n = 11", "n = 28", "n = 17", "n = 24"),
                          nDOR = c("n = 10", "n = 9", "n = 1", "n = 11", "n = 11", "n = 28", "n = 17", "n = 24"),
                          LocNum = c("1", "2", "2", "3", "3", "4", "5", "6"))
prac_labels <- prac_labels %>% 
  mutate(LocSp2 = fct_reorder(LocSp2, LocNum))

#creating dataframe to use when plotting mean vline
means <- graph_data_2 %>% 
  dplyr::group_by(LocSp2) %>% 
  dplyr::summarise(meanSL = mean(SL),
            meanRGR1 = mean(RGR1, na.rm = T),
            meanRGR2 = mean(RGR2, na.rm = T),
            meanVERT = mean(VERT),
            meanDOR = mean(DOR_XRAY))
join_labels <- data.frame(LocSp2 = c("FA", "NBA", "NBM", "EA", "EM", "VM", "SAM", "PM"),
                          Species = c("australis", "australis", "melanochir", "australis", "melanochir", "melanochir", "melanochir", "melanochir"),
                          LocNum = c("1", "2", "2", "3", "3", "4", "5", "6"))
means <- left_join(means, join_labels, by = "LocSp2")
means <- means %>% 
  mutate(LocSp2 = fct_reorder(LocSp2, LocNum))

```

#5. Final histograms

NOTE: these histograms were then imported into Adobe Illustrator to edit site labels and compile separate graphs into a single image for publication

```{r}
RGR1plot_annotated <- RGR1plot +
  geom_text(x = 38, y = 5, aes(label = nRGR1), size = 3.6, data = prac_labels) +
  annotate("text", x = 30, y = 5, label = " ") +
  geom_vline(data = means, aes(xintercept=meanRGR1, colour = Species), linetype = "dashed", size = 1) +
  theme(axis.text = element_text(colour = "black"),
        axis.title = element_text(size = 14),
        legend.text = element_text(face = "italic", size = 12),
        legend.title = element_text(size = 13))
RGR2plot_annotated <- RGR2plot +
  geom_text(x = 32.5, y = 5, aes(label = nRGR2), size = 3.6, data = prac_labels) +
  annotate("text", x = 32.5, y = 5, label = " ") +
  geom_vline(data = means, aes(xintercept=meanRGR2, colour = Species), linetype = "dashed", size = 1) +
  theme(axis.text = element_text(colour = "black"),
        axis.title = element_text(size = 14),
        legend.text = element_text(face = "italic", size = 12),
        legend.title = element_text(size = 13))
VERTplot_annotated <- VERTplot +
  geom_text(x = 59, y = 10, aes(label = nVERT), size = 3.6, data = prac_labels) +
  annotate("text", x = 59, y = 10, label = " ") +
  geom_vline(data = means, aes(xintercept=meanVERT, colour = Species), linetype = "dashed", size = 1) +
  theme(axis.text = element_text(colour = "black"),
        axis.title = element_text(size = 14),
        legend.text = element_text(face = "italic", size = 12),
        legend.title = element_text(size = 13))
SLplot_annotated <- Sizeplot +
  geom_text(x = 300, y = 5, aes(label = nSL), size = 3.6, data = prac_labels) +
  annotate("text", x = 300, y = 5, label = " ") +
  geom_vline(data = means, aes(xintercept=meanSL, colour = Species), linetype = "dashed", size = 1) +
  theme(axis.text = element_text(colour = "black"),
        axis.title = element_text(size = 14),
        legend.text = element_text(face = "italic", size = 12),
        legend.title = element_text(size = 13))
DORplot_annotated <- DORplot +
  geom_text(x = 15, y = 12, aes(label = nDOR), size = 3.6, data = prac_labels) +
  annotate("text", x = 17, y = 15, label = " ") +
  geom_vline(data = means, aes(xintercept=meanDOR, colour = Species), linetype = "dashed", size = 1) +
  theme(axis.text = element_text(colour = "black"),
        axis.title = element_text(size = 14),
        legend.text = element_text(face = "italic", size = 12),
        legend.title = element_text(size = 13))

#saving plots
ggsave(RGR1plot_annotated, file = "~/Data Analysis/results/meristics/RGR1_histogram.jpg")
ggsave(RGR2plot_annotated, file = "~/Data Analysis/results/meristics/RGR2_histogram.jpg")
ggsave(VERTplot_annotated, file = "~/Data Analysis/results/meristics/VERT_histogram.jpg")
ggsave(SLplot_annotated, file = "~/Data Analysis/results/meristics/SL_histogram.jpg")
ggsave(DORplot_annotated, file = "~/Data Analysis/results/meristics/DOR_histogram.jpg")

```


#***Morphometrics***

#1. Setup

```{r}
#reloading raw dataset
raw_dat <- read.csv("~/Data Analysis/data/Whole_Fish_Raw_Measurements.csv", header = TRUE)

#loading packages
library(tidyverse)
library(ggpubr)
library(rstatix)
library(broom)

```

#2. Cleaning dataset

```{r}
# removing unknown species and damaged specimens
raw_dat_2 <- raw_dat %>% 
  filter(Species != "?",
         Species != "NA",
         Sample_ID != "E30")

# setting species as a factor
raw_dat_2$Species <- factor(raw_dat_2$Species)

# calculating "postorbital" length, i.e. head length - upper jaw length - preorbital length - orbital length
raw_dat_2 <- raw_dat_2 %>% 
  mutate(POSTORB = HDL - UJL - PREORB - ORB)

# creating new long-standardized dataframe
logdata <- raw_dat_2 %>% 
  dplyr::select(2:4, 14, 16:28, 31) %>%
  mutate_if(is.numeric, log)

```

#3. Testing ANCOVA assumptions

This code was run individually for all morphometric variables.
"UJL" is put in the following example.

```{r}
#1. Linearity - scatterplots were created and saved in "data standardization" step 3

#2. Homogeneity of regression slopes 
logdata %>% 
  anova_test(UJL ~ Species*SL)

#3. normality of residuals (want shapiro test to not be significant)
testmodel <- lm(UJL ~ SL*Species, data = postorb_data)
model.metrics <- augment(testmodel) %>% 
  select(-.hat, -.sigma, -.fitted, -.se.fit)
head(model.metrics)
shapiro_test(model.metrics$.resid)
hist(logdata$UJL)

#4. Homogeneity of variance (also want to not be signif)
model.metrics %>% levene_test(.resid ~ Species)

# 5. outliers
model.metrics %>% 
  filter(abs(.std.resid) > 3) %>%
  as.data.frame()
# creating seperate dataframes with outliers removed for variables where outliers were identified 
hdl_data <- logdata %>% 
  filter(Sample_ID != "SA19")
abase_data <- logdata %>% 
  filter(Sample_ID != "E24",
         Sample_ID != "E29")
p1p2_data <- logdata %>% 
  filter(Sample_ID != "E33")
orb_data <- logdata %>% 
  filter(Sample_ID != "E32",
         Sample_ID != "E40")
preorb_data <- logdata %>% 
  filter(Sample_ID != "E27",
         Sample_ID != "WA20")
postorb_data <- logdata %>% 
  filter (Sample_ID != "V13")

```

#4. ANCOVA
All results were saved as plain text before manual input into a table 

```{r}
ujl.res.aov <- logdata %>% 
  anova_test(UJL ~ SL + Species)
get_anova_table(ujl.res.aov)

ujw.res.aov <- logdata %>% 
  anova_test(UJW ~ SL + Species)
get_anova_table(ujw.res.aov)

hdl.res.aov <- hdl_data %>% 
  anova_test(HDL ~ SL + Species)
get_anova_table(hdl.res.aov)

p1l.res.aov <- logdata %>% 
  anova_test(P1L ~ SL + Species)
get_anova_table(p1l.res.aov)

dbase.res.aov <- logdata %>% 
  anova_test(DBASE ~ SL + Species)
get_anova_table(dbase.res.aov)

abase.res.aov <- abase_data %>% 
  anova_test(ABASE ~ SL + Species)
get_anova_table(abase.res.aov)

p1p2.res.aov <- p1p2_data %>% 
  anova_test(P1.P2 ~ SL + Species)
get_anova_table(p1p2.res.aov)

p2c.res.aov <- logdata %>% 
  anova_test(P2.C ~ SL + Species)
get_anova_table(p2c.res.aov)

bdp1.res.aov <- logdata %>% 
  anova_test(BD.P1 ~ SL + Species)
get_anova_table(bdp1.res.aov)

bdp2.res.aov <- logdata %>% 
  anova_test(BD.P2 ~ SL + Species)
get_anova_table(bdp2.res.aov)

orb.res.aov <- orb_data %>% 
  anova_test(ORB ~ SL + Species)
get_anova_table(orb.res.aov)

preorb.res.aov <- preorb_data %>% 
  anova_test(PREORB ~ SL + Species)
get_anova_table(preorb.res.aov)

postorb.res.aov <- postorb_data %>% 
  anova_test(POSTORB ~ SL + Species)
get_anova_table(postorb.res.aov)

```

#5. Violin plots
#a) Setup

```{r}
#loading required packages
library(ggplot2)
library(dplyr)
library(tidyverse)

#loading dataset - this contains the residuals from the "Data Standardization"section, as well as grouping variables (location number and location + species)
boxdata <- read.csv("~/2021-Docs/2021 Uni/Publishing/All R Work/data/violin_plot_data.csv", header = T)

# creating a theme to be used on all violin plots
theme_violin <- theme(panel.grid = element_blank(),
                      panel.background = element_blank(),
                      legend.position = "bottom",
                      strip.placement = "inside",
                      strip.text.y.left = element_text(angle=0, hjust=1),  
                      strip.background = element_rect(colour = "white", fill = "white"),
                      axis.text.x = element_text(colour = "black"),
                      axis.title = element_text(size = 14),
                      legend.text = element_text(face = "italic", size = 12),
                      legend.title = element_text(size = 13))
```

#b) Cleaning dataset

```{r}
# Editing the grouping variables so they can be used to facet plots
boxdata <- boxdata %>% 
  mutate(LocSp = fct_reorder(LocSp, Location_Num),
         Location = fct_reorder(Location, Location_Num))

#creating shorter names for values in plots
join_labels <- data.frame(LocSp2 = c("FA", "NBA", "NBM", "EA", "EM", "VM", "SAM", "PM"),
                          LocSp = c("Forster-australis", "Nelson Bay-australis", "Nelson Bay-melanochir", 
                                    "Eden-australis", "Eden-melanochir", "Corner Inlet, Victoria-melanochir",
                                    "Adelaide, South Australia-melanochir", "Perth, Western Australia-melanochir"))

boxdata <- left_join(boxdata, join_labels, by = "LocSp")
boxdata <- boxdata %>% 
  mutate(LocSp2 = fct_reorder(LocSp2, Location_Num))

# creating seperate datasets with outliers removed for variables with outliers identified through ANCOVA
preorb_data <- boxdata %>% 
  filter(Sample_ID != "E27",
         Sample_ID != "WA20")
orb_data <- boxdata %>%
  filter(Sample_ID != "E32",
         Sample_ID != "E40")
hdl_data <- boxdata %>% 
  filter(Sample_ID != "SA19")

```

#c) creating and saving plots

NOTE: these violin plots were then imported into Adobe Illustrator to edit site labels, manually add sample numbers, and compile separate graphs into a single image for publication

```{r}
# Violin plots split by location

final_preorb <- ggplot(preorb_data, aes(x=PREORB, y=LocSp2, colour = Species, fill = Species)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.2, fill = "white", colour = "black") +
  scale_colour_manual(values = c("darkorange1", "deepskyblue1")) +
  scale_fill_manual(values = c("#FFCC99", "#99CCFF")) +
  scale_y_discrete(limits = rev(levels(preorb_data$LocSp2)))+
  labs(x="Preorbital Length") +
  theme_violin


final_orb <- ggplot(orb_data, aes(x=ORB, y=LocSp2, colour = Species, fill = Species)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.2, fill = "white", colour = "black") +
  scale_colour_manual(values = c("darkorange1", "deepskyblue1")) +
  scale_fill_manual(values = c("#FFCC99", "#99CCFF")) +
  scale_y_discrete(limits = rev(levels(orb_data$LocSp2)))+
  labs(x="Orbital Length") +
  theme_violin

final_hdl <- ggplot(hdl_data, aes(x=HDL, y=LocSp2, colour = Species, fill = Species)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.2, fill = "white", colour = "black") +
  scale_colour_manual(values = c("darkorange1", "deepskyblue1")) +
  scale_fill_manual(values = c("#FFCC99", "#99CCFF")) +
  scale_y_discrete(limits = rev(levels(hdl_data$LocSp2)))+
  labs(x="Head Length") +
  theme_violin

final_ujl <- ggplot(boxdata, aes(x=UJL, y=LocSp2, fill = Species, colour = Species)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.2, fill = "white", colour = "black") +
  scale_colour_manual(values = c("darkorange1", "deepskyblue1")) +
  scale_fill_manual(values = c("#FFCC99", "#99CCFF")) +
  scale_y_discrete(limits = rev(levels(boxdata$LocSp2)))+
  labs(x="Upper Jaw Length") +
  theme_violin

final_p1l <- ggplot(boxdata, aes(x=P1L, y=LocSp2, fill = Species, colour = Species)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.2, fill = "white", colour = "black") +
  scale_colour_manual(values = c("darkorange1", "deepskyblue1")) +
  scale_fill_manual(values = c("#FFCC99", "#99CCFF")) +
  scale_y_discrete(limits = rev(levels(boxdata$LocSp2)))+
  labs(x="Pectoral Fin Length") +
  theme_violin

# Violin plots split by species

sp_final_preorb <- ggplot(preorb_data, aes(x=PREORB, y=Species, colour = Species, fill = Species)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.2, fill = "white", colour = "black") +
  scale_colour_manual(values = c("darkorange1", "deepskyblue1")) +
  scale_fill_manual(values = c("#FFCC99", "#99CCFF")) +
  scale_y_discrete(limits = rev)+
  labs(x="Preorbital Length") +
  theme_violin

sp_final_orb <- ggplot(orb_data, aes(x=ORB, y=Species, colour = Species, fill = Species)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.2, fill = "white", colour = "black") +
  scale_colour_manual(values = c("darkorange1", "deepskyblue1")) +
  scale_fill_manual(values = c("#FFCC99", "#99CCFF")) +
  scale_y_discrete(limits = rev)+
  labs(x="Orbital Length") +
  theme_violin

sp_final_hdl <- ggplot(hdl_data, aes(x=HDL, y=Species, colour = Species, fill = Species)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.2, fill = "white", colour = "black") +
  scale_colour_manual(values = c("darkorange1", "deepskyblue1")) +
  scale_fill_manual(values = c("#FFCC99", "#99CCFF")) +
  scale_y_discrete(limits = rev)+
  labs(x="Head Length") +
  theme_violin

sp_final_ujl <- ggplot(boxdata, aes(x=UJL, y=Species, fill = Species, colour = Species)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.2, fill = "white", colour = "black") +
  scale_colour_manual(values = c("darkorange1", "deepskyblue1")) +
  scale_fill_manual(values = c("#FFCC99", "#99CCFF")) +
  scale_y_discrete(limits = rev)+
  labs(x="Upper Jaw Length") +
  theme_violin

sp_final_p1l <- ggplot(boxdata, aes(x=P1L, y=Species, fill = Species, colour = Species)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.2, fill = "white", colour = "black") +
  scale_colour_manual(values = c("darkorange1", "deepskyblue1")) +
  scale_fill_manual(values = c("#FFCC99", "#99CCFF")) +
  scale_y_discrete(limits = rev)+
  labs(x="Pectoral Fin Length") +
  theme_violin

#saving all plots 
ggsave("~/Data Analysis/results/morphometrics/loc_preorb.jpg", plot = final_preorb)
ggsave("~/Data Analysis/results/morphometrics/loc_orb.jpg", plot = final_orb)
ggsave("~/Data Analysis/results/morphometrics/loc_hdl.jpg", plot = final_hdl)
ggsave("~/Data Analysis/results/morphometrics/loc_ujl.jpg", plot = final_ujl)
ggsave("~/Data Analysis/results/morphometrics/loc_p1l.jpg", plot = final_p1l)
ggsave("~/Data Analysis/results/morphometrics/spp_preorb.jpg", plot = sp_final_preorb)
ggsave("~/Data Analysis/results/morphometrics/spp_orb.jpg", plot = sp_final_orb)
ggsave("~/Data Analysis/results/morphometrics/spp_hdl.jpg", plot = sp_final_hdl)
ggsave("~/Data Analysis/results/morphometrics/spp_ujl.jpg", plot = sp_final_ujl)
ggsave("~/Data Analysis/results/morphometrics/spp_p1l.jpg", plot = sp_final_p1l)
```




