---
title: "Otolith Data Analysis"
output: html_notebook
---

Author: Indiana Riley

This markdown follows the instructions outlined in Libungan & Pálsson, 2015. See this paper for full instruction and explanation:
Libungan, L. A., & Pálsson, S. (2015). ShapeR: An R package to study otolith shape variation among fish populations. PLoS ONE, 10(3). https://doi.org/10.1371/journal.pone.0121102


#1. Setup

```{r}
library(shapeR)

FISH <- read.csv("C:/Users/Indy/Desktop/ShapeAnalysis5/FISH.csv", header = T)

myshape = shapeR("C:/Users/Indy/Desktop/ShapeAnalysis5", "FISH.csv")

```

#2. Creating outline reconstructions
(only need to run this once and then save)

```{r}
myshape = detect.outline(myshape,threshold=0.2,write.outline.w.org=TRUE)

save(myshape, file = "newshapefile.RData")

load("newshapefile.RData")

myshape = remove.outline(myshape, "AU", "TG11_Ha6173")
myshape = remove.outline(myshape, "AU", "TG18_Ha6180")
myshape = remove.outline(myshape, "AU", "S9_Ha6211")
myshape = remove.outline(myshape, "AU", "K7_Ha5826")
myshape = remove.outline(myshape, "AU", "U10_Ha5870")
myshape = remove.outline(myshape, "AU", "NB40")
myshape = remove.outline(myshape, "ME", "V01")
myshape = remove.outline(myshape, "ME", "V06")
myshape = remove.outline(myshape, "ME", "V09")
myshape = remove.outline(myshape, "ME", "SA16")
myshape = remove.outline(myshape, "ME", "SA17")
myshape = remove.outline(myshape, "ME", "SA20")

myshape = smoothout(myshape, n = 100)

```

#3. Generating shape coefficients

```{r}
myshape = generateShapeCoefficients(myshape)

myshape = enrich.master.list(myshape)

getMeasurements(myshape)

# the code below extracts the summary statistics for each otolith as a csv

measurements <- getMeasurements(myshape)
masterlist <- getMasterlist(myshape)
masterlist <- masterlist[1:20]
write.csv(masterlist, file = "otolith_measurements.csv", row.names = FALSE)

```

#4. Graphs

```{r}
tapply(getMeasurements(myshape)$otolith.area, getMasterlist(myshape)$species, mean)
tapply(getMeasurements(myshape)$otolith.area, getMasterlist(myshape)$state, mean)

plotWaveletShape(myshape, "species", show.angle = TRUE, lwd = 2, lty = 1, col = c("darkorange1", "deepskyblue1"))
plotWaveletShape(myshape, "state", show.angle = TRUE, lwd = 2, lty = 1, col = 4:9)


myshape = stdCoefs(myshape, classes = "species", "length_cm", bonferroni = FALSE)
est.list = estimate.outline.reconstruction(myshape)
outline.reconstruction.plot(est.list, max.num.harmonics = 15)
par(mfrow = c(1,1))
plotWavelet(myshape, level = 5, class.name = "species", useStdcoef = TRUE)

```

#5. Analysis

Note that random forest and permanova were not used in Libungan & Pálsson, 2015

```{r}
# capscale analysis of variance

library(vegan)
cap.res = capscale(getStdWavelet(myshape) ~ getMasterlist(myshape)$species)
anova(cap.res, by = "terms", step = 1000)

# Random forest 
library(randomForest)

stdw = getStdWavelet(myshape)
species = factor(getMasterlist(myshape)$species)
dd = data.frame(stdw = stdw,species = species)

out_E <- randomForest(formula = as.factor(species)~., data=dd, ntrees=5000)

out_E
importance(out_E)
par(mfrow=c(1,1))
plot(out_E)

# PERMANOVA
## Checking assumptions
dist_euc <- vegdist(getStdWavelet(myshape), method = "euclidean")
Disp <- betadisper(dist_euc, getMasterlist(myshape)$species)

anova(Disp) #(P = 0.1089) # Therefore assumption for PERMANOVA is met

## Running test
dist_euc <- vegdist(getStdWavelet(myshape), method = "euclidean")
adonis(dist_euc ~ getMasterlist(myshape)$species)

```





